Supplementary information: In-silico molecular enrichment and clearance of the human intracranial space

### 2025-06-09 paper/main.tex
revision I from Marius III

Now, to examine how CSF pulsatility affects the tracer enrichment patterns,
we consider three variations of the baseline: (i) no CSF production,
(ii) reduced pulsatility and thus decreased dispersion ($0.1 \times
R$), and (iii) higher pulsatility with increased dispersion ($10
\times R$). Without CSF production, transport is considerably delayed
(\Cref{fig:csf}N, O). Even after $12$ hours, tracer remains in the
subtentorial regions around the cerebellum, the brain stem and in the
surrounding CSF. Interestingly, the lack of CSF production instead
allows tracer to travel upwards through the ventricular system,
reaching the third ventricle after around $12$ hours. On the other
hand, if the CSF pulsatility is reduced, we observe rapid transport
towards the upper convexity of the cranium, as in the baseline model,
but the tracer spreads exclusively within the anterior regions
(\Cref{fig:csf}P). This feature can be attributed to the CSF flow
bifurcation posterior to the ambient cistern (\Cref{fig:csf}D):
without sufficient diffusion, the tracer is unable to cross into the
posterior SAS. Indeed, with higher dispersion, the tracer moves
through the quadrigeminal cistern and further upwards into the
longitudinal fissure, with also enrichment of the cerebellum
(\Cref{fig:csf}Q).

>>>>

Now, to examine how CSF pulsatility affects the tracer enrichment patterns,
we consider three variations of the baseline: (i) no CSF production,
(ii) reduced pulsatility and thus decreased dispersion, and (iii) higher pulsatility with increased dispersion. We computed the reduced and increased pulsatility scenarios by halving and doubling the inflow rates associated with both cardiac and respiratory pulsations, which reduced the dispersion enhancement factors by factor of 0.25 and 4, respectively. Note that behaviour directly reflects the nature of our model: The Stokes flow is linear in its inflow boundary conditions, and the diffusion enhancement factor quadratic in the predicted pressure gradient.
Without CSF production, transport is considerably delayed
(\Cref{fig:csf}N, O). Even after $12$ hours, tracer remains in the
subtentorial regions around the cerebellum, the brain stem and in the
surrounding CSF. Interestingly, the lack of CSF production instead
allows tracer to travel upwards through the ventricular system,
reaching the third ventricle after around $12$ hours. On the other
hand, if the CSF pulsatility is reduced, we observe rapid transport
towards the upper convexity of the cranium, as in the baseline model,
but the tracer spreads exclusively within the anterior regions
(\Cref{fig:csf}P). This feature can be attributed to the CSF flow
bifurcation posterior to the ambient cistern (\Cref{fig:csf}D):
without sufficient diffusion, the tracer is unable to cross into the
posterior SAS. Indeed, with higher dispersion, the tracer moves
through the quadrigeminal cistern and further upwards into the
longitudinal fissure, with also enrichment of the cerebellum
(\Cref{fig:csf}Q).

### 2025-06-09 paper/supplementary.tex
revision I from Marius III

% Dispersion predictions
\paragraph{Dispersion in the SAS, ventricular system and PVS}
This pulsatile flow of CSF in the SAS, ventricular system and PVSs
leads to an increase in effective solute
diffusivity~\cite{stockman2007effect, hettiarachchi2011effect,
  asgari2016glymphatic, sharp2019dispersion, ray2021quantitative} via
a process known as Taylor dispersion \cite{taylor1953dispersion,
  watson1983diffusion}. Previous estimates of the magnitude of this
effect in the CSF spaces vary significantly: from an enhancement
factor of 0.05--1 in periarterial spaces surrounding penetrating
arteries~\cite{asgari2016glymphatic, troyetsky2021dispersion}, to
5--100 in the spinal subarachnoid space~\cite{stockman2007effect,
  hettiarachchi2011effect, sharp2019dispersion}, and up to more than
$10\,000$ in surface periarterial spaces~\cite{ray2021quantitative,
  sharp2019dispersion}. The large variability can (at least partly) be
attributed to methodological differences; e.g. different assumptions
on the medium, domain width, pressure differences and/or fluid
velocities, the diversity of CSF flow characteristics, as well as a
high likelihood of spatial variations. Hornkjøl et
al~\cite{hornkjol2022csf} consider model variations with constant
dispersion factors from 1 up to 1000, and indicate that a value of 10
gives the better agreement with the clinically observed
enrichment. Our spatially-varying estimates of the dispersion
enhancement factors $R_c, R_r$ (with $D = (1 + R_c + R_r) D^{\rm
  Gad}$) range from 0 to 200 for the cardiac contribution $R_c$ and 0
to 320 for the respiratory contribution $R_r$; and is thus compatible
within the previously reported spectrum.

>>>>

% Dispersion predictions
\paragraph{Dispersion in the SAS, ventricular system and PVS}
This pulsatile flow of CSF in the SAS, ventricular system and PVSs
leads to an increase in effective solute
diffusivity~\cite{stockman2007effect, hettiarachchi2011effect,
  asgari2016glymphatic, sharp2019dispersion, ray2021quantitative} via
a process known as Taylor dispersion \cite{taylor1953dispersion,
  watson1983diffusion}. Previous estimates of the magnitude of this
effect in the CSF spaces vary significantly: from an enhancement
factor of 0.05--1 in periarterial spaces surrounding penetrating
arteries~\cite{asgari2016glymphatic, troyetsky2021dispersion}, to
5--100 in the spinal subarachnoid space~\cite{stockman2007effect,
  hettiarachchi2011effect, sharp2019dispersion}, and up to more than
$10\,000$ in surface periarterial spaces~\cite{ray2021quantitative,
  sharp2019dispersion}.
Additionally, high-resolution MRI measurements by Hirschler et al.~\cite{hirschler2024region} revealed that CSF mobility in regions with high pulsatility is approximately tenfold greater than the self-diffusion of water.
The large variability can (at least partly) be
attributed to methodological differences; e.g. different assumptions
on the medium, domain width, pressure differences and/or fluid
velocities, the diversity of CSF flow characteristics, as well as a
high likelihood of spatial variations. Hornkjøl et
al~\cite{hornkjol2022csf} consider model variations with constant
dispersion factors from 1 up to 1000, and indicate that a value of 10
gives the better agreement with the clinically observed
enrichment. Our spatially-varying estimates of the dispersion
enhancement factors $R_c, R_r$ (with $D = (1 + R_c + R_r) D^{\rm
  Gad}$) range from 0 to 200 for the cardiac contribution $R_c$ and 0
to 320 for the respiratory contribution $R_r$; and is thus compatible
within the previously reported spectrum.

Investigating the link between cardiac pulsatility and the resulting diffusion enhancement factor, we compute for $R$ different levels of cardiac pulsatility. We reduce the inflow boundary conditions of $6\,$ml/s (across the SAS outer surface) and $0.31\,$ml/s (across the lateral ventricle surface) by $25\,\%$, $50\,\%$, $75\,\%$ and $100\,\%$ to model lower cardiac pulsatility, reaveling that $R$ scales quadratically with the reduction of inflow (\cref{fig:R_comparison}). This behaviour directly reflects the nature of our model - the Stokes flow is linear in its boundary conditions, and the diffusion enhancement factor quadratic in the predicted pressure gradient - and simultaneously underscores the substantial role of pulsatility in shaping intracranial transport.

\begin{figure}
    \centering
    \includegraphics[width=1.0\linewidth]{paper/figures/cardiac_R_comparison.png}
    \caption{Comparison of cardicac-driven diffusion enhancement factors for different levels of cardiac pulsatility. From left to right: The diffusion enhancement factor $R_c$ for $25\,\%$, $50\,\%$, $75\,\%$ and $100\,\%$ of the inflow boundary conditions of $6\,$ml/s and $0.31\,$ml/s across the SAS outer surface and the lateral ventricle surface, respectively.}
    \label{fig:R_comparison}
\end{figure}
